---
title: "Data Processing & Analysis: Dow Jones Industrial Average Index"
author: "Handerson Coq"
date: '2022-06-05'
output: html_document
---

### Context

We are interested in understanding stock price movements over the years to see if we can identify some patterns.

  * The dataset for this project is retrieved from this [Kaggle](https://www.kaggle.com/datasets/deeplytics/dow-stock-data?select=dow_historic_2000_2020) project. It has data from 2000 to 2020.

  * We have already performed data cleaning with this dataset in SQL Server and are familiar with it. Therefore, steps such as verifying whether the data is complete, checking for nulls and missing values, will not be necessary here.

  * The list of DJIA companies is retrieved from [Investopedia](https://www.investopedia.com/ask/answers/050115/what-does-dow-jones-industrial-average-measure.asp#:~:text=The%20Dow%20Jones%20Industrial%20Average%20(DJIA)%20measures%20the%20daily%20price,the%20transportation%20and%20utility%20industries.)


### Data processing (cleaning)

```{r libraries, echo=TRUE, message=FALSE, warning=FALSE}

# Importing libraries

library(tidyverse)
library(dplyr)
library(readxl)
library(janitor)
library(skimr)

```

```{r Stock Data, echo=TRUE, message=FALSE, warning=FALSE}

# Importing DJIA stock data

dow_jones_stocks <- read.csv("dow_historic_2000_2020.csv")

head(dow_jones_stocks)

```

```{r DJIA Companies, echo=TRUE, message=FALSE, warning=FALSE}

# Importing the DJIA company list

dow_jones_companies <- read_xlsx("dow_jones_companies.xlsx")

dow_jones_companies

```

```{r Stock Data Cleaning, echo=TRUE, message=FALSE, warning=FALSE}

# Stock data processing & cleaning

# Checking the datatypes and column names

str(dow_jones_stocks)

# The date column is in char, changing to date

dow_jones_stocks$date <- as.Date(dow_jones_stocks$date)

# Verifying that the date column is in correct format and has no duplicates

str(dow_jones_stocks$date)

length(unique(dow_jones_stocks$date)) # works if returns 5284

```

```{r DJIA Comp. Clean, echo=TRUE, message=FALSE, warning=FALSE}

# DJIA Companies data processing & cleaning

# Cleaning the column names and rename the symbol column to stock

colnames(dow_jones_companies)

dj_comp_clean <- clean_names(dow_jones_companies)
  
dj_comp_clean <- rename(dj_comp_clean, stock = symbol)

```

```{r Join data, echo=TRUE, message=FALSE, warning=FALSE}

# Joining the stock dataset to the DJIA company dataset

dj_stocks_clean <- dow_jones_stocks %>% 
  inner_join(dj_comp_clean, by = "stock")

# We will remove the year_added and industry_type from the new dataset

dj_stocks <- dj_stocks_clean[,-c(12, 13)]

head(dj_stocks)
tail(dj_stocks)

# Skim through the data

skim_without_charts(dj_stocks)

```

### Exploratory Analysis

```{r Join data, echo=TRUE, message=FALSE, warning=FALSE}


```

